---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE}
library(dplyr,warn.conflicts = FALSE)
library(ggplot2,quietly=TRUE,warn.conflicts = FALSE)
library(xtable,quietly=TRUE,warn.conflicts = FALSE)
library(gridExtra,quietly=TRUE,warn.conflicts = FALSE)

unzip("./activity.zip")
data <- read.csv("./activity.csv")

```

## What is mean total number of steps taken per day (NA data removed)?

<br> Below is a historgram of the steps taken each day:
```{r echo=TRUE}
#Remove NA data
data_no_na <- na.omit(data)
#Summarize data
grp_by_StepsPerDay <- group_by(data_no_na,date)
summary <- summarize(grp_by_StepsPerDay,total_steps=sum(steps),mean_steps=mean(steps),median_steps=median(steps))
#Plot data
ggplot(summary, aes(x=total_steps)) + geom_histogram() + scale_y_continuous(breaks=seq(1,10,1)) + xlab("Mean Steps/Day") + ggtitle("Histogram of Mean Steps Per Day")
```

<br> Below is the mean and media of the total number of steps taken each day (NA data removed):
```{r showtable,results='asis'} 

#Print data
xt <- xtable(summary)
print(xt,type = "html")

```

## What is the average daily activity pattern  (NA data removed)?
```{r echo=TRUE}
grp_by_MeanStepsPerInterval <- group_by(data_no_na,interval) 
summary2 <- summarize(grp_by_MeanStepsPerInterval,interval_mean=mean(steps))

ggplot(summary2, aes(interval, interval_mean)) + geom_line() + xlab("Interval") + ylab("Mean Steps")
```
<br>

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
filter(summary2,interval_mean==max(summary2$interval_mean))$interval
```

## Imputing missing values

<br>Total number of missing values in the dataset:
```{r echo=TRUE}
sum(is.na(data$steps))
```
<br> <br> Below is a historgram of the steps taken each day with NAs replaced by interval means:
```{r echo=TRUE}

#Default rows with NA for steps to interval mean
mergedData <- merge(data,summary2)
mergedData$steps <- ifelse(is.na(mergedData$steps),mergedData$interval_mean,mergedData$steps)

#Create histogram of total steps per day
grp_by_StepsPerDay2 <- group_by(mergedData,date)
summary3 <- summarize(grp_by_StepsPerDay2,total_steps=sum(steps),mean_steps=mean(steps),median_steps=median(steps))
ggplot(summary3, aes(x=total_steps)) + geom_histogram() + scale_y_continuous(breaks=seq(1,10,1)) + xlab("Mean Steps/Day") + ggtitle("Histogram of Mean Steps Per Day")

```

<br> Below is the mean and media of the total number of steps taken each day with NAs replaced by interval means:

```{r showtable2,results='asis'}
#Print data
xt2 <- xtable(summary3)
print(xt2,type = "html")

```

### There does not seem to be a major impact of replaced NA data with interval means

## Are there differences in activity patterns between weekdays and weekends?

<BR>Below are two graphs that capture average steps per interval for weekends and weekdays.

```{r echo=TRUE}

mergedData$dayofweek <- weekdays(as.Date(mergedData$date))
weekends <- mergedData[mergedData$dayofweek %in% c("Saturday","Sunday"),]
weekdays <- mergedData[!mergedData$dayofweek %in% c("Saturday","Sunday"),]

grp_by_MeanStepsPerIntervalWE <- group_by(weekends,interval)
grp_by_MeanStepsPerIntervalWD <- group_by(weekdays,interval)

summaryWE <- summarize(grp_by_MeanStepsPerIntervalWE,interval_mean=mean(steps))
summaryWD <- summarize(grp_by_MeanStepsPerIntervalWD,interval_mean=mean(steps))
pWE <- ggplot(summaryWE, aes(interval, interval_mean)) + geom_line() + xlab("Weekend Interval") + ylab("Mean Steps")
pWD <- ggplot(summaryWD, aes(interval, interval_mean)) + geom_line() + xlab("Weekday Interval") + ylab("Mean Steps")

grid.arrange(pWE, pWD)

```

### There does seem to be difference between weekend and weekday activity. The average of Weekend intervals ramp up and remain at a more constant level of activity relative to weekdays which seems to spike and then level out at a lower average steps per interval. 
